---
title: "Informal Sensitivity Analysis of Almond Yield Profit"
author: "Bailey Jørgensen and Michelle Yiv"
format: html
editor: visual
---

## Load libraries

```{r}
library(tidyverse)
library(here)
library(purrr)
```

## Read in data and source functions

```{r}
# Read in data for use in the analysis
daily_data <- read.table("clim.txt", header = TRUE, sep = " ", stringsAsFactors = FALSE)

# Source the function
source("almond_profit.R")
```

## Run model

```{r}
# Use default parameters to run model
almond_result <- almond_profit(daily_data)
```

## Informal Sensitivity analysis


```{r}
# Define deviation and baseline values for production cost and almond price/ton

deviation <- 0.12
nsamples = 100
base_production_cost <- 4000
base_price <- 5000

# Generate samples using uniform distribution. Assumed as we don't know variation

production_cost_samples <- runif(
  min = base_production_cost - deviation * base_production_cost,
  max = base_production_cost + deviation * base_production_cost, n = nsamples
)


almond_price_samples <- runif(
  min = base_price - deviation * base_price,
  max = base_price + deviation * base_price, n = nsamples
)

# Now place samples in df
parms <- cbind.data.frame(production_cost = production_cost_samples, price = almond_price_samples)
```

```{r}
# Apply function using pmap
results <- parms %>% 
  pmap(function(production_cost, price) {
    almond_profit(
      price = price,
      production_cost = production_cost
    )
  })
```

```{r}
# Combine all result data frames into one long data frame
results_df <- map_df(results, bind_rows, .id = "run") %>%
  mutate(run = as.integer(run))

# Add run ID to parms
parms <- parms %>%
  mutate(run = row_number())

# Join parameter values to the results
sensitivity <- left_join(results_df, parms, by = "run")
```


## Graphs

- one that shows yield anomaly for each year, accounting for uncertainty in the parameters
- one that show how yield anomaly in general varies with your parameters

```{r}
# Yield anomaly by year
ggplot(sensitivity, aes(x = year, 
                          y = yield_anomaly, 
                          )) +
  geom_line() + 
  labs(title = "Yield Anomaly Over Years",
       x = "Year", y = "Yield Anomaly (tons/acre)") + 
  theme_minimal() +
  theme(legend.position = "none") 

```
```{r}
# Yield anomaly by year
ggplot(sensitivity, aes(x = year, 
                          y = profit, 
                          )) +
  geom_line() + 
  labs(title = "Profit by year",
       x = "Year", y = " Profit($/acre)") + 
  theme_minimal() +
  theme(legend.position = "none") 
```

```{r}
# Get unique values for each run 
unique_vals <- sensitivity %>%
  select(run, mean_profit, production_cost) %>%
  distinct()

# Now plot
ggplot(unique_vals, aes(x = production_cost, y = mean_profit)) +
  geom_point(size = 2, color = "pink") +
  labs(x = "Production Cost ($/acre)",
       y = "Mean Profit ($/acre)",
       title = "Mean Profit vs Production Cost") +
  theme_minimal()

```
```{r}
# Same plot as above but as boxplot
ggplot(unique_vals, aes(x = production_cost, y = mean_profit)) +
  geom_boxplot() +
  labs(x = "Production Cost ($/acre)",
       y = "Mean Profit ($/acre)",
       title = "Mean Profit vs Production Cost") +
  theme_minimal()
```


```{r}
# Plot price sensitivity
ggplot(sensitivity, aes(x = price, y = mean_profit)) +
  geom_point() +
  labs(y = "Profit (dollar/acre)", 
       x = "Price (dollar/ton)") +
  theme_classic()
```
```{r}
# Get unique values for each run 
unique_vals <- sensitivity %>%
  select(run, mean_yield_anomaly, production_cost) %>%
  distinct()

# Plot yield anomaly with production cost
ggplot(unique_vals, aes(x = mean_yield_anomaly, y = production_cost)) +
  geom_boxplot() +
  labs(y = "Production cost ($/acre)", 
       x = "Mean Yield Anomaly") +
  theme_classic()
```
```{r}
# Get unique values for each run 
unique_vals <- sensitivity %>%
  select(run, mean_yield_anomaly, price) %>%
  distinct()

# Plot yield anomaly with production cost
ggplot(unique_vals, aes(x = mean_yield_anomaly, y = price)) +
  geom_boxplot() +
  labs(y = "Price", 
       x = "Mean Yield Anomaly") +
  theme_classic()
```

## Interpretation

Looking at yield anomalies over the years, there is a large spike in 1995, with the next biggest spikes being similar in size in 2005 and 2008. These patterns are also seen when looking at profits over the years, although the scales are different. This indicates that these years were different from normal yields and saw favorable harvests. When looking at mean profit and production cost, the boxplot looks fairly equal, which can indicate that profit is not as sensitive to production cost.
Now looking at price sensitivity (price vs mean profit), this plot makes sense as price of almonds/ton being sold increases, so does profit. This is a positive linear relationship, assuming that other parameters do not vary significantly. 
Finally, we look at how our mean yield anomalies vary with our chosen parameters. The production cost boxplot has its lower quartile smaller than its upper quartile, meaning that there is positive or right skew. This means that most of the production costs fall in the lower range. The price boxplot appears to be fairly equal. The price per ton of almonds is relatively stable in this dataset, although that is likely due to how the sensitivity analysis was performed. Additionally, this also suggests that mean yield anomalies do not influence almond price per ton a lot. 
